Calculating Module Enrichment and Visualizing Data on Large-scale Molecular Maps with the R packages \pkg{ACSNMineR} and \pkg{RNaviCell}

### 2016-01-07 jss/article.tex
[skip ci] started to write breast cancer example.

>>>>

\subsection{Analysis of breast cancer expression data}
In a study published in 2008, Schmidt and colleagues analyzed gene expression
patterns of 200 breast cancer patients not treated by systemic therapy after
surgery using discovery approach to reveal additional prognostic motifs
\cite{schmidt2008humoral}. Estrogen receptor (ER) expression and proliferative
activity of breast carcinomas are well-known and described prognostic markers.
Patients with ER-positive carcinomas have a better prognosis than those with
ER-negative carcinomas, and rapidly proliferating carcinomas have an adverse
prognosis. Knowledge about the molecular mechanisms involved in the
processes of estrogen-dependent tumor growth and proliferative activity has led
to the successful development of therapeutic concepts, such as  antiendocrine and
cytotoxic chemotherapy.

The dataset corresponding to this study is available as a
Bioconductor package. The code shown below is creating a list of differentially
expressed genes

\begin{verbatim}

library(breastCancerMAINZ)
library(Biobase)
library(limma)
library(ACSNMineR)
library(hgu133a.db)
library(RNaviCell)

data(mainz)
eset <- exprs(mainz)
pdat <- pData(mainz)
design <- model.matrix(~factor(pdat$er == '1'))
lmFit(eset, design) -> fit
eBayes(fit) -> ebayes
toptable(ebayes, coef=2,n=25000)->tt
#which(abs(tt$logFC) >=1 & tt$adj < 0.05) -> sel
which(tt$adj < 0.05) -> sel
rownames(tt[sel,]) -> probe_list
mget(probe_list, env=hgu133aSYMBOL) -> slist
slist <- as.character(slist)

enrichment(slist) -> res

#apply(eset[probe_list,pdat$er == 1],1,mean)->er_plus_mean
#apply(eset[probe_list,pdat$er == 0],1,mean)->er_minus_mean
#cbind(er_plus_mean, er_minus_mean)-> data_er
#rownames(data_er) <- slist

#n$map_url <- "http://acsn.curie.fr/navicell/maps/apoptosis/master/index.php"
#n$importDatatable("mRNA expression data", "exp", data_er)

apply(eset[probe_list,pdat$er == 0],1,mean) -> er_minus_mean
names(er_minus_mean) <- slist
er_minus_mean <- as.matrix(er_minus_mean)
colnames(er_minus_mean) <- c('exp')

n <- NaviCell()
n$proxy_url <- "https://acsn.curie.fr/cgi-bin/nv_proxy.php"
n$map_url <- "https://acsn.curie.fr/navicell/maps/acsn/master/index.php"

n$launchBrowser()
n$importDatatable("mRNA expression data", "GBM_exp", er_minus_mean)
n$heatmapEditorSelectSample('0','exp')
n$heatmapEditorSelectDatatable('0','GBM_exp')
n$heatmapEditorApply()

# analysis mtsig signature file

#mt <- format_from_gmt('/Users/eric/Downloads/c2.cp.v5.0.symbols.gmt')
#unique(slist)->l
#> length(l)
#[1] 5179
#> head(l)
#[1] "ESR1"   "CA12"   "GATA3"  "TBC1D9" "FOXA1"  "IL6ST" 
#> enrichment(l, maps=mt)
